### 2017-02-04 response.tex
revised

\documentclass{paper}

\title{Author's response to the review of No.IJPRAI-D-16-00429}

\begin{document}

\begin{center}
{\Large \textbf{Author's response to the review of No.IJPRAI-D-16-00429\\[20pt]}}
\end{center}

\textbf{Dear Editors and Reviewers:\\}

We would like to thank the reviewers and associate editors for their
valuable and constructive comments and suggestions. Based on these
comments and suggestions, we revised the paper, and the main revisions
are as follows.\\

\begin{center}
\textbf{Revisions Based on the Comments of Reviewer 1}\\
\end{center}

\textbf{1.} The construction of AMT used a breadth-first search
strategy, why does not use a depth-first strategy instead? Is there
some inherent reasons? Please give an explanation.\\[7pt]
\noindent \textbf{Response:} There is no essential distinction between
the breadth-first strategy and the depth-first strategy in
constructing the AMT. If the breadth-first strategy is adopted, the
AMT will be built level by level; on the other hand, if the
depth-first strategy is used, the AMT will be built branch by
branch. But regardless of the strategies used, the final built AMTs
are the same. We choose the breadth-first strategy to build the AMT
level by level, and we have added this explanation to Section 5.1.\\

\textbf{2.} Is the hash function used in the "Hash Table" node
structure the same as in other algorithms such as WM and
Pre-filter+AC?\\[7pt]
\noindent \textbf{Response:} No, they are not the same. The hash
functions used by the WM and Pre-filter+AC algorithms apply
essentially to short strings which consist no more than 3 characters,
whereas the hash function used in the \emph{hash node} must be able to
process longer strings with more than 3 characters. Therefore, we
choose the \emph{shift-add-xor} hash family for the hash node to
process long
strings.\\

\textbf{3.} The chosen of the parameters is very important for
building the tree nodes. Please give a more detailed instruction to
the chosen of the parameters in section 6.1.\\[7pt]
\noindent \textbf{Response:} For String Arrays containing no more than
\textbf{4} strings, the simple linear search is faster than binary
search due to the complexity of binary search. The load factor of the
Hash Table is set to \textbf{0.5}, which is a good balance between
time and space. The parameters about the hash function are chosen from
its original paper, which are claimed to be very efficient. We have
tested that once $ndp > 100$, the hash table structure becomes more
efficient than the String Array structure. All these detailed
explanations for parameter
selections have been added to section 6.1.\\

\textbf{4.} The captions and some table items in table 2, 3 and 4 are
not very clear, please revise them.\\[7pt]
\noindent \textbf{Response:} The captions of Table 2, 3 and 4 are
revised. The \emph{Sing Char} and \emph{Single String} items in these
tables are modified to \emph{Map 1} and \emph{Single String Array}, respectively.\\

\begin{center}
\textbf{Revisions Based on the Comments of Reviewer 2}\\
\end{center}

\textbf{1.} Is the hash function used in the \emph{Hash Table} node structure
the same as in the filter module?\\[7pt]
\noindent \textbf{Response:} No, they are not the same. The hash
function used in the \emph{Hash Table} node comes from the
\emph{shift-add-xor} hash family which is able to process long
strings, whereas the hash function used in the filter module is
usually a direct map which maps a short string to its decimal
representation.\\

\textbf{2.} Please explain how the two parameters $N$ and $k$ are chosen in the
filter module.\\[7pt]
\noindent \textbf{Response:} If $lsp \leq 5$, set $k = 2$; otherwise,
set $k=3$. $N$ is set to be $|\Sigma|^k$. We have added this comments
to Section 4.\\

\textbf{3.} The authors choose to use a breadth-first strategy to
build the AMT, is it possible to use a depth-first strategy instead?
\noindent \textbf{Response:} Absolutely yes. Regardless of which
strategies are used, the final constructed AMTs are the same.

\textbf{4.} The example (Fig.2) in the filter module does not clearly
show the interaction with the verification module. Please add some
detailed illustrations.
\noindent \textbf{Response:} Every time the filter module detects a
potential matched position in the text (i.e., the last bit of
$MB~\&~QB$ is 1), the verification module is invoked to check if there
exists patterns starting at that position, and in the mean while, the
matching window stops at that position, once the verification
finishes, the filter module moves the matching window to the next
position. We have added this detailed procedure to Section 4.\\

\begin{center}
\textbf{Revisions Based on the Comments of Reviewer 3}
\end{center}

There is no comments from reviewer 3.\\

\begin{center}
\textbf{Revisions Based on the Comments of Reviewer 4}
\end{center}

\textbf{1.} The novelty or fresh new ideas is not very clear, does not
have enough comparisons with other methods in the literature.\\[7pt]
\noindent \textbf{Response:} The major contributions of this paper
are: 1. A novel data structure called Adaptive Matching Tree (AMT) is
proposed to store the string pattern set efficiently. 2. A fast engine
based on AMT is proposed, which is able to process large-scale string
pattern sets. In addition, we have added another comparison with
a recent algorithm to the experiment part of our paper.\\

\textbf{2.} What is its relation to PR \& AI ? (Pattern Recognition \&
Artificial Intelligence) Should explain it more clearly.\\[7pt]
\noindent \textbf{Response:} The term \emph{pattern} has various
meanings in various fields. In this paper, the term \emph{pattern} is
specified to \emph{string pattern}, and accordingly \emph{pattern
  recognition} is specified to \emph{string pattern searching}, which
is one of the
oldest and fundamental problems in computer science.\\

\textbf{3.} Need to add more recent related publications in the
reference list including , for examples 2015, 2016 PR \& AI subfield.\\[7pt]
\noindent \textbf{Response:} We have already added several recent
published (in 2015 or 2016) references
about PR \& AI to our paper.\\

\textbf{4.} Compare your work with others in the literature,
advantages vs disadvantages, in depth and width.\\[7pt]
\noindent \textbf{Response:} We have added another comparison in our
experiment to a recent published multi-string pattern matching
algorithm called \textsf{Split} proposed by \emph{Kim, HyunJin AND
  Choi, Kang-Il AND Choi, Sang-Il. A Memory-Efficient Deterministic
  Finite Automaton-Based Bit-Split String Matching Scheme Using
  Pattern Uniqueness in Deep Packet
  Inspection. PLOS ONE. vol.10. pp.1-24. 2015.}\\

\textbf{5.} Should use real images with noisy and testing samples set
size should be large enough to show your results are indeed
convincing,
reliable, and meaningful.\\[7pt]
\noindent \textbf{Response:} Unfortunately, our method only focuses
on string patterns, so it does not apply to images.\\

\textbf{6.} This paper is not very well written. Its English
presentation is not good enough. Should consult English technical
writing expert to help.\\[7pt]
\noindent \textbf{Response:} We have carefully revised the paper again
to improve its presentation, so we hope it is more readable and easy
to understand now.

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End: