### 2017-02-04 A_Fast_Engine_for_Multi_string_Pattern_Matching(revised).tex
revised

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Trim Size: 9.75in x 6.5in
%% Text Area: 8in (include Runningheads) x 5in
%% ws-ijprai.cls   :   8-8-2014
%% Class file to use with ws-ijprai.tex written in Latex2E.
%% The content, structure, format and layout of this style file is the
%% property of World Scientific Publishing Co. Pte. Ltd.
%% Copyright 2014 by World Scientific Publishing Co.
%% All rights are reserved.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%

\documentclass{ws-ijprai}
\usepackage{enumerate}

\usepackage{graphicx}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{multirow}
\usepackage{amsmath}
\usepackage{color}
\usepackage{pgfplots}

\newcommand{\SWITCH}[1]{\STATE \textbf{switch} (#1)}
\newcommand{\ENDPWITCH}{\STATE \textbf{end switch}}
\newcommand{\CASE}[1]{\STATE \textbf{case} #1\textbf{:} \begin{ALC@g}}
\newcommand{\ENDCASE}{\end{ALC@g}}
\newcommand{\CASELINE}[1]{\STATE \textbf{case} #1\textbf{:} }
\newcommand{\DEFAULT}{\STATE \textbf{default:} \begin{ALC@g}}
\newcommand{\ENDDEFAULT}{\end{ALC@g}}
\newcommand{\DEFAULTLINE}[1]{\STATE \textbf{default:} }
\newtheorem{mydef}{Definition}
\newtheorem{mylm}{Proposition}

\renewcommand{\algorithmicrequire}{\textbf{Input:}}
\renewcommand{\algorithmicensure}{\textbf{Output:}}

\begin{document}

\markboth{Zhan Peng et al.}{A Fast Engine for Multi-string Pattern Matching}

%%%%%%%%%%%%%%%%%%%%% Publisher's area please ignore %%%%%%%%%%%%%%%
%
\catchline{}{}{}{}{}
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\title{A Fast Engine for Multi-string Pattern Matching}

\author{Zhan Peng, Yuping Wang\footnote{Corresponding author}~, Wei Yue}

\address{$^1$School of Computer Science and Technology\\
  Xidian University\\
  Xi'an, Shaanxi, 710071, China\\
   $^*$ywang@xidian.edu.cn
}

\maketitle

%begin{history}
%received{(Day Month Year)}
%revised{(Day Month Year)}
%accepted{(Day Month Year)}
%comby{(xxxxxxxxxx)}
%end{history}

\begin{abstract}
  Multi-string matching is a core technique searching a text string
  for all occurrences of some string patterns. It is widely used in
  many applications. However, as the number of string patterns
  increases, most of the existing algorithms suffer from two issues:
  the long matching time, and the high memory consumption. To address
  these issues, in this paper, a fast matching engine is proposed for
  large-scale string matching problems. Our engine includes a filter
  module and a verification module. The filter module is based on
  several bitmaps which are response for quickly filtering out the
  invalid positions in the text, while for each potential matched
  position, the verification module confirms true pattern
  occurrence. In particular, we design a compact data structure called
  Adaptive Matching Tree (AMT) for the verification module, in which
  each tree node only saves some pattern fragments of the whole
  pattern set and the inner structure of each tree node is chosen
  adaptively according to the features of the corresponding pattern
  fragments. This makes the engine time and space efficient. The
  experiments indicate that, our matching engine performs better than
  the compared algorithms, especially for large pattern sets.
\end{abstract}

\keywords{String Matching; Pattern Matching; Dictionary Matching}

% due to the effectiveness of the filter and the compactness of the AMT
% whether there are patterns starting at that position

\section{Introduction}
\label{sec:introduction}

String matching (SM) has been historically one of the key areas of
computer science. According to the number of strings to be matched
simultaneously, string matching technologies can be classified into
single string matching (SSM) and multi-string matching (MSM), and this
paper mainly focuses on the latter.

The MSM is a primitive but critical operation used in many
applications such as pattern recognition \cite{Yan2016},
\cite{Xiao2016}, ontology alignments \cite{Xue2015} \cite{Xue2016},
text classification \cite{Tang2015} \cite{Zhang2016}, intrusion detection system
\cite{Kim2015}, etc. In general, it seeks all positions in a text
string that match any of the string patterns in a given pattern
set. An efficient algorithm for MSM only needs to scan the text once
and finds all patterns concurrently. Traditionally, the existing
algorithms include two major phases, i.e. the \textsf{preprocessing
phase} and the \textsf{matching phase}. In the \textsf{preprocessing
phase}, a suitable data structure is constructed to store all the
string patterns, and then the \textsf{matching phase} scans and
compares the given text with the data structure to find all
occurrences of every pattern in the text. It is obvious that the time
and space efficiency of the data structure built in the preprocessing
phase can have a significant impact on the performance of the
algorithm in the matching phase.

Nowadays, the number of patterns to be matched becomes huge in some
applications, for instance, the database of a modern
anti-virus/intrusion-detection system can easily contain millions
number of virus signatures/rules. However, the existing algorithms are
not suitable for such applications with very large pattern sets due to
two main reasons: 1) The poor scalability of the data structures used;
2) The space inefficiency of the data structure used, which means
these algorithms can not process large pattern sets with limited
memory. Moreover, the existing algorithms are sensitive to the length
of the patterns, particularly the length of the shortest pattern. Many
algorithms can not efficiently deal with pattern set containing very
short pattern due to the poor robustness of their data
structures. Therefore, it is very valuable to design space-efficient
data structures with good scalability and robustness to support
applications requiring large-scale string matching.

In this paper, we propose a multi-string matching engine with good
robustness for large string pattern sets. Similarly to the existing
efficient approaches, our approach consists of the preprocessing phase
and the matching phase. In the preprocessing phase, the matching
engine is constructed based on the pattern set. Then in the matching
phase, the engine takes the text string as input, and searches the
occurrences of each pattern in the text.

The proposed engine consists of a filter module and a verification
module. The function of the filter module is to find the potential
matched positions in the text. Once such a position is found, the
verification module confirms whether there are patterns starting at
that position. The filter adopted is proposed in \cite{Lee2013}, which
is based on several bitmaps and a group of simple bit-wised
"\emph{and}" and "\emph{shift}" operations. In the matching phase, the
engine queries the bitmaps to check the validity of the current
position: if the current position is a potential matched position, the
verification module is invoked; otherwise, the engine skips over this
position as well as the next few positions as many as possible. In
fact, the filter is proved to be optimal in the sense that it is able
to utilizing all previous query results, and thus can skip as many
invalid positions as possible. For the verification module, we design
a compact tree-like data structure called Adaptive Matching Tree
(\textsf{AMT}) based on the pattern set. Specifically, the whole
pattern set is gradually partitioned into several small parts first,
and then the pattern fragments of each small part is stored in a
separate node of the AMT for subsequent matching. The critical feature
of AMT is that, the inner structure of each tree node is changed
adaptively according to the features of the pattern fragments, and
this adaptivity guarantees a high matching speed as well as a low
memory consumption for each tree node, as a result, the whole AMT
would be both time and space efficient for verifying in the matching
phase.

The rest of this paper is organized as follows: Section
\ref{sec:related works} reviews some related works in MSM. Section
\ref{sec:notations} lists the notations and terminology used in this
paper. Section \ref{sec:filter} describes the filter module. Section
\ref{sec:verification} presents the verification module. Section
\ref{sec:experiments} presents the performance evaluations. Section
\ref{sec:conclusion} concludes the paper and proposes some open
problems.

\section{Related Works}
\label{sec:related works}

Because of the importance of PM, several efficient algorithms have
been proposed in the past decades. Well-known algorithms for PM
include Knuth-Morris-Pratt (\textsf{KMP}) \cite{Knuth1977},
Boyer-Moore (\textsf{BM}) \cite{Boyer1977}, Wu-Manber (\textsf{WM})
\cite{Wu1994}, and Aho-Corasick (\textsf{AC}) \cite{Aho1975}. The
\textsf{KMP} and \textsf{BM} algorithms are appropriate for SPM, but
not suitable for MPM. The \textsf{AC} algorithm is a generalization of
the \textsf{KMP} algorithm by preprocessing the patterns and building
a deterministic finite automaton (DFA) that can match multiple
patterns simultaneously. It is the first algorithm for MPM that has a
linear time complexity. However, the \textsf{AC} algorithm is
alphabet-sensitive and requires a huge amount of memory to build the
DFA, which limits its application in large-scale pattern matching.

To reduce the memory requirement of \textsf{AC}, Tuck \cite{Tuck2004}
proposed an algorithm called \textsf{Bitmapped AC}. This algorithm
uses the bitmap data structure to replace the conventional child
pointers in each state. In addition, it uses a path compression
technique which can merge successive single-child states into one
state. It is reported that, the compression optimization of
\textsf{Bitmapped AC} results in a 50 times reduction in memory
consumption over the default \textsf{AC} algorithm. However, since
traversal from node to node requires checking a bit in a bitmap and
then performing a sum up to 256 prior bits in the bitmap, the average
performance of this algorithm is not high.  Another algorithm to
reduce the space requirement is the \textsf{Compressed AC} algorithm
proposed by \cite{Bremler2011}. In addition to use a similar path
compression method, this algorithm also develops a technique that can
eliminate the leaf states. The experiment shows that
\textsf{Compressed AC} can reduce at most $60\%$ memory requirement
compared with the traditionally \textsf{AC} algorithm. But it only
works well for problems with a few pattern sets. Lee \cite{Lee2013}
proposed an algorithm called \textsf{Pre-filter+AC}. As the name
shows, this algorithm consists of a pre-filter and a verification
engine. The function of the pre-filter is to query data structures
built from patterns to find the starting positions of potential
pattern occurrences. Once a suspicious starting position is found, the
verification engine confirms true pattern occurrence. The pre-filter
uses a bit vector, called master bitmap, with simple bitwise-AND and
shift operations to accumulate query results. While the verification
engine, which is a modification of the AC automaton, checks all
candidate patterns simultaneously rather than sequentially. However,
as the author mentioned, the pre-filter is suitable for patterns of
moderate or large lengths. For short patterns, its throughput
degrades. Recently, Kim \cite{Kim2015} proposed a memory-efficient
algorithm \textsf{Split} for deep packet inspection system, which
consists of multiple finite-state machine.

The \textsf{WM} algorithm is another efficient algorithm for MPM which
is known to be faster than \textsf{AC} in a typical matching. The
\textsf{WM} algorithm is an adaptation of the \textsf{BM} algorithm to
multiple patterns. It uses a similar ``shift idea'' of the \textsf{BM}
algorithm, but the shift of \textsf{WM} is not based on a single
character as the \textsf{BM} does, instead, it is based on a
$B$-$gram$ that is a character block consists of $B$ characters ($B$
is usually set to 2 or 3 in practice). In a typical search, its time
complexity is $O(B \cdot n/lsp)$, where $lsp$ is the length of the
shortest pattern in the pattern set. Therefore, the performance of WM
is seriously affected by $lsp$.

To improve the \textsf{WM} algorithm, Zhou \cite{Zhou2007} proposed an
algorithm called \textsf{MDH}. Instead of using the first $m$
characters as the signature of each pattern, \textsf{MDH} adopts a
heuristic strategy to select the optimum $m$ consecutive characters as
the signature of each pattern. Similarly, Zhan \cite{Zhan2014}
proposed a different way to search for the optimal signatures of the
patterns, additionally, this algorithm designs an index structure to
reduce the time for searching the candidate patterns in the Hash
Table. These algorithms can improve the performance of \textsf{WM} for
common patterns but they do not solve the inefficiency caused by short
patterns. To address this problem, Zhang \cite{Zhang2009} proposed the
High Concurrence WM algorithm (\textsf{HCWM}), which separates the
patterns whose length is less than 4 from the pattern set. For these
short patterns, the algorithm establishes independent data structures
and uses different matching routines. This method can reduce the
effect of short patterns, since short patterns are matched
independently and concurrently. Another algorithm for this problem is
the \textsf{$L^{+1}$-MWM} proposed by Choi \cite{Choi2011}, which
minimizes the performance degradation caused by the short patterns by
appending characters to these patterns. It is reported that, the
\textsf{$L^{+1}$-MWM} improves the performance of WM by as mush as
$20\%$ in average, moreover, when $lsp < 5$ the \textsf{$L^{+1}$-MWM}
gains a $38.87\%$ enhancement.

For large scale pattern matching, which becomes more and more
important in big data context, Le \cite{Le2013} presented a
Memory-efficient Architecture for large-scale String Matching
(\textsf{MASM}) based on a pipelined binary search tree. It uses a
technique called ``leaf-attaching'' to compress the given pattern
set. The compressed pattern set are then transformed to a pipelined
binary tree which will be used in the matching phase.  It is reported
that, the \textsf{MASM} achieves a memory efficiency (defined as the
ratio of the amount of the required storage in bytes and the size of
the pattern set in number of characters) of 0.56 for the Rogets
dictionary and 1.32 for the Snort rule set. Moraru \cite{Moraru2013}
also presented a memory-efficient and cache-optimized algorithm for
large pattern sets. This algorithm builds upon the Rabin-Karp
\cite{Karp1987} SPM algorithm and incorporates a new
\textsf{feed-forward} bloom filter which takes into account the memory
hierarchy of modern computers. This algorithm is also well suited for
implementation on GPUs which enables the matching to be done in
parallel.

There are also studies that exploited the capacity of hardware to
accelerate the proposed MPM engines. Agarwal \cite{Agarwal2013}
proposed a hardware architecture which enables high throughout for
Information Extraction applications. Instead of using the common DFA
based approach, this architecture employs a novel hashing based
scheme. The DFA based approaches (such as \textsf{AC}) typically
process one character every cycle, while the proposed hash based
scheme can process a string token of several characters every cycle,
thus achieves higher throughout than the DFA based approaches. Zhang
\cite{Zhang2015} proposed an GPU-based parallel algorithm
\textsf{G-PEBF} which uses the Extended Bloom Filter (EBF). It divides
the pattern set into $N$ subsets where the lengths of the patterns in
each subsets are the same. Then it constructs an EBF for each subset
and uses $N$ threads to simultaneously process the subsets in parallel
on the GPU. The performance of \textsf{G-PEFB} highly relies on the
number of threads used during matching, and it is also not applicable
for pattern sets containing very short patterns.

Nowadays, some variants of the classical MPM problem have been
proposed. Tomohiro \cite{I2015} introduced a variant of MPM, where the
pattern set is given in a compressed form that can be represented by a
straight line program (SLP). For a given SLP-compressed pattern set of
size $n$ and height $h$, which represents $m$ patterns of total length
$N$, they present an $O(n^2log^N)$-size variant of AC automaton that
recognizes all occurrences of the patterns in $O(h + m)$ running time
per character. Khancome \cite{Khancome2013} proposed an algorithm for
the dynamic MPM problem. This algorithm uses the \textsf{inverted
  lists} data structure to allow the pattern set to be updated
dynamically in an optimal time. Amir \cite{Amir2015} introduced an
algorithm for the gapped MPM problem, where each pattern in the
pattern set is a sequence of sub-patterns separated by bounded
sequences of do not cares. Neuburger \cite{Neuburger2012} presented
the first efficient algorithm that operates in small space for the
2-dimensional MPM problem.

\section{Notations}
\label{sec:notations}

Let $\Sigma$ be an \textsf{alphabet} consisting of a finite number of
character symbols. (In this paper we will only focus on the case that
$\Sigma$ is the ASCII character set, i.e. $|\Sigma| = 256$ and each
character takes one byte memory.) Given the alphabet $\Sigma$, a
\textsf{string} as well as its \textsf{substrings} over
$\Sigma$ can be defined as follows:\\
\\
\textbf{Definition 1.} A \textsf{string} over $\Sigma$ is a sequence
consisting of a finite number of characters from $\Sigma$. A
length-$n$ string is represented by $S = c_1c_2..c_n$, where the
$i$-th character of $S$ is: $S[i] = c_i \in \Sigma$
$(1 \leq i \leq n)$. The length of $S$ is denoted by $|S|$. A
\textsf{substring} of $S$ is a sequence consisting of any consecutive
characters of $S$. The substring of $S$, which starts at
position $i$ and has length $len$, is denoted by $S[i,\,len]$.\\
\\
Given a string $S$, its \textsf{prefix} and \textsf{suffix},
which are both special substrings of $S$, are defined as follows:\\
\\
\textbf{Definition 2.} The length-$m$ \textsf{prefix} of a string $S$,
which is denoted by $S^{+m}$, is the substring that consists of the
first $m$ characters of $S$, i.e. $S^{+m}=S[1,\,m]$.  While the
\textsf{suffix} of $S$ that starts at position $m+1$, is the substring
left after removing the length-$m$ prefix of $S$. It is denoted by
$S^{-m}$, i.e. $S^{-m} = S[m+1,\,|S|-m]$. (Particularly, a string $S$
can be regarded as a prefix/suffix of itself, i.e. $S=S^{+|S|}=S^{-0}$.)\\
\\
Finally, the MSM problem can be defined as follows: \\
\\
\textbf{Definition 3.} Given a text string $T$ and a pattern set
$P=\{p_1,\,p_2,\,\dots,\,p_k\}$, where $T$ and $p_i$
$(1 \leq i \leq k)$ are all strings over $\Sigma$. The multi-string
matching problem is to find all occurrences of every pattern of $P$ in
$T$, more formally, to build the result set
$R = \{(i,\, p_j)\;|\; p_j \in P\; and\,\; T[i,\,|p_j|]=p_j\}$.

\section{The Filter Module}
\label{sec:filter}

As mentioned before, our engine consists of a filter module and a
verification module. In this section, we introduce the former, while
in the next section the latter will be illustrated.

The filter we adopted in our engine is proposed by \cite{Lee2013},
which is an improvement of the filter in the WM algorithm. Given the
pattern set, as in the WM algorithm, only the length-$lsp$ prefix of
each pattern are considered in constructing the filter, where $lsp$ is
the \emph{length of the shortest pattern} in the pattern set. Similar
to WM, the filtering is based on character blocks (a character block
is merely a short string consisting of only a few characters) rather
than single character. Assuming that each block consists of $k$
characters, the filter will include $lsp-k+1$ membership query bitmaps
denoted as $B_1$, $B_2$, \dots, and $B_{lsp-k+1}$, where each bitmap
is a vector of $N$ bits. (If $lsp \leq 5$, set $k = 2$; otherwise, set
$k=3$. $N$ is set to be $|\Sigma|^k$). Let $B_j[i]$,
$0 \leq i \leq N - 1$, $1 \leq j \leq lsp-k+1$, be the $i$-th bit of
the $j$-th query bitmap, and $B_j[i]$ is set to 1 iff there exists a
pattern $p_m$ such that $hash(p_m[j,k]) = i$ ($p_m[j,k]$ is a
$k$-symbol block starting at position $j$ of $p_m$), where $hash$ is a
hash function that maps a $k$-symbol block to an integer in the range
of $[0, N-1]$.

\begin{figure}[htbp]
  \centering
  \includegraphics[height=2in, width=2in]{./eps/filter}
  \caption{The filter constructed from three digital patterns.}
  \label{fig:filter}
\end{figure}

Figure \ref{fig:filter} shows an example of the filter, assume that
alphabet $\Sigma = \{0, 1, 2, \dots, 9\}$ and pattern set
$P = \{p_1,\; p_2,\; p_3\}$, where $p_1 = 04648753$, $p_2 = 30692$,
and $p_3 = 614621$.  Obviously, $lsp = |p_2|= 5$. Assume that $k = 2$,
$N = 100$, and the hash function is the identity mapping. For such a
setting, there are $lsp - k + 1 = 4$ query bitmaps, i.e. $B_1$, $B_2$,
$B_3$, $B_4$, and in particular, we have $B_1[i] = 1$ iff $i$ = 4, 30,
or 61; $B_2[i] = 1$ iff $i$ = 6, 14, or 46; $B_3[i] = 1$ iff $i$ = 46,
64, or 69; and $B_4[i] = 1$ iff $i$ = 48, 62, or 92.

Once the query bitmaps have been constructed, they can be used to
filter the positions of the input text string. During filtering, a
sliding window $W$ of length $lsp$ is adopted to select a segment of
the text string $T$. Initially, $W$ is aligned with the beginning of
$T$, so that the substring of $T$ contained in $W$ is $T[1,lsp]$.
Generally, assume that $W$ is moved to the $i$-th position of $T$, and
the substring of $T$ contained in $W$ is $T[i,lsp]$. Then the last
$k$-symbol block of $T[i,lsp]$, which is $T[i+lsp-k, k]$, is used to
query all the bitmaps. Let single bit $qb_j$ be the report of $B_j$
($1 \leq j \leq lsp - k + 1$), and $qb_j=1$ iff
$B_j[hash(T[i+lsp-k,k])] = 1$. Let bitmap
$QB = qb_1qb_2 \dots qb_{lsp-k+1}$ to denote the current query result
and another \emph{master bitmap} $MB = mb_1mb_2 \dots mb_{lsp-k+1}$ to
accumulate the previous query results and act as the current state of
the filter. Initially, all the bits of $MB$ are set to 1. After
fetching a query result $QB$, the master bitmap is updated by
$MB = MB \; \& \; QB$, where \& is the bit-wised ``and''
operation. For the updated $MB$, if $mb_{lsp-k+1} = 1$, position $i$
is considered as a potential matched position and the verification
module is invoked. Otherwise, the sliding window $W$ is advanced by
$lsp-k+1$ positions if all the bits of $MB$ are 0's, or $lsp-k+1-r$
positions if $mb_r=1$ and $mb_i=0$, for $r < i \leq lsp-k$. The
positions skipped by $W$ are filtered out. If $W$ is determined to be
advanced by $h$ positions, $MB$ is right-shifted by $h$ bits and
filled with 1's for the holes left by the shift. Particularly, it is
proved in \cite{Lee2013} that the filter using master bitmap and
simple bit-wised ``and'' and ``shift'' operations is optimal in the
sense that it is able to utilize all previous query results, which
means the filter can filter out as many positions as possible in the
text.

\begin{figure}[htbp]
  \centering
  \includegraphics[height=2.5in, width=2.5in]{./eps/filter_match}
  \caption{Filter the positions in the text using the filter.}
  \label{fig:f_match}
\end{figure}

As an example (see Figure \ref{fig:f_match}), consider the query bitmaps
constructed above and assume that $T=23764614621$.  Initially, the
matching window $W$ contains 23764 and $MB = 1111$.  The last 2-symbol
block of $W$, i.e. 64, is used for the first query, and the query
result is $QB=0010$. Since $MB\; \&\; QB = 0010$, $W$ is advanced by 1
position (which means the first position of $T$ is filtered out), and
then $MB$ is updated as $1001$ by right shift. For the second
position, the substring contained in $W$ is 37646. This time, block 46
is used to query the bitmaps and the query result is $QB=0110$. Since
$MB\; \& \; QB=0000$, $W$ moves forward by 4 positions to position 6
(which means positions $2 \sim 5$ are filled out), and $MB$ is updated
as 1111. For position 6, $W$ contains 61462 and block 62 is used for
querying and fetches $QB = 0001$. Since $MB\; \& \; QB = 0001$,
position 6 is a potential matched position, and thus, the verification
module is invoked. After checking, the pattern $p_3=614621$ is
detected at position 6. Then, $W$ is advanced by 4 positions to
position 10. However, since the length of the remaining input text in
$W$, i.e. 21, is smaller than $lsp$, which means there can not be any
pattern starting at that position, the whole scanning procedure is
terminated.

\section{The verification module}
\label{sec:verification}

The verification module is the core of the whole matching engine,
since it determines the worst case performance of the engine. The
verification module in our engine is based on a new data structure
called \emph{Adaptive Matching Tree} (AMT) which is a modification of
the classical \emph{trie} structure. Like the \emph{trie}, but in a
more flexible and space efficient way, the AMT enables a fast
membership query which decides whether a given target string exists in
the pattern set. In the next subsections, we will introduce the
construction of the AMT and some further optimization techniques.

\subsection{AMT Construction}
\label{subsec:amt}

The AMT is constructed from the pattern set, and in general, this
construction includes five major steps:

\begin{enumerate}
\item Create a suffix set $SF$, and put each pattern in the pattern
  set into $SF$, where each pattern is regarded as a suffix of itself.
\item Compute the \emph{length of the shortest suffix} ($lss$) in
  $SF$. For each suffix in $SF$, remove its length-$lss$ prefix, and
  all the removed prefixes form a prefix set $PF$. Discard the
  repetitions in $PF$, and calculate $|PF|$ which refers to the
  \emph{number of distinct prefixes} ($ndp$).
\item Create a tree \emph{node} to hold the prefixes in $PF$. The
  \emph{node} is actually an index, in which each prefix serves as a
  key. The inner structure of the \emph{node} is constructed
  adaptively according to $ndp$ and $lss$ that have been computed in
  step 2.
\item Divide $SF$ by putting together all the suffixes with the same
  prefix removed, to form a sub-suffix-set. Associate each
  sub-suffix-set with the corresponding prefix in the tree
  \emph{node}.
\item For each newly created sub-suffix-set in step 3, repeat the same
  procedure from step 2.
\end{enumerate}

\begin{figure}[htbp]
  \centering
  \includegraphics[width=\textwidth]{./eps/AMT}
  \caption{Constructing the AMT from a pattern set. Each arrow
    in the AMT indicates a child pointer.}
  \label{fig:AMT}
\end{figure}

As an example, Figure \ref{fig:AMT} shows the construction of AMT from
a pattern set $P = \{p_1,\, p_2,\, \dots,\, p_{13}\}$. Note that,
since each pattern is a suffix of itself, the whole pattern set $P$
can be regarded as a suffix set $SF$.  First, since $p_6$ is the
shortest pattern(suffix) in $P$, we have $lss = |p_6| = 2$. Then the
length-2 prefix of each pattern(suffix) is removed, and the removed
prefixes form a prefix set:
$PF = \{p_1^{+2} = aa,\; p_2^{+2} = aa,\; \dots,\; p_{13}^{+2} =
cc\}$.  After discarding the repetitions in $PF$, only 3 distinct
prefixes are left: $PF = \{aa,\, bb,\, cc\}$, and we have
$ndp = |PF| = 3$.

Then the first tree node, which is actually the root of AMT, is
created to hold the prefixes in $PF$. Each prefix serves as a
\emph{key} in the root and has a pointer (initialized to $NULL$) which
will be used to associate with the child node in the future. As will
be seen latter, the structure of root is selected adaptively according
to two parameters: $lss=2$ and $ndp=3$.

After creation of the root, the suffixes left in $P$ are grouped based
on their lost 2-symbol prefixes: those suffixes having the same
prefixes removed are grouped together to form a sub-suffix-set. In
order to record the correspondence between the sub-suffix-set and the
lost prefix, the tuple (\emph{node.key,\; sub-suffix-set}) is used:
the suffixes in the \emph{sub-suffix-set} have the same prefix
\emph{key} removed, where the \emph{key} is held as a key in the newly
created \emph{node}. Using this notation, there are three tuples
created after grouping the suffixes left in $P$:
$tp_1 = (root.aa,\; SF_1=\{p_1^{-2},\, p_2^{-2},\, p_3^{-2},\,
p_4^{-2},\, p_5^{-2},\, p_7^{-2}\})$,\,
$tp_2 = (root.bb,\; SF_2=\{p_8^{-2},\, p_9^{-2},\, p_{10}^{-2}\})$ and
$tp_3 = (root.cc,\; SF_3=\{p_{11}^{-2},\, p_{12}^{-2},\,
p_{13}^{-2}\})$. This indicates that $P$ is divided into three
sub-suffix-sets: $SF_1$, $SF_2$, $SF_3$, and each sub-suffix-set
corresponds to a prefix held as a key in the root (it also means that
the root may have three child nodes). Note that, the shortest
suffix(pattern) $p_6$ has disappeared in the sub-suffix-sets, since
after removing the length-2 prefix of $p_6$, there is nothing left.

The same procedure is repeated on each of the three created
sub-suffix-sets. As a further example, $SF_1$ of $tp_1$ is
processed. First, since $lss = |p_7^{-2}| = 6$, the length-$6$ prefix
of each suffix in $SF_1$ is removed to form a prefix set:
$PF_1 = \{(p_1^{-2})^{+6},\, (p_2^{-2})^{+6},\, (p_3^{-2})^{+6},\,
(p_4^{-2})^{+6},\, (p_5^{-2})^{+6},\, (p_7^{-2})^{+6}\}$. After
discarding the duplicates in $PF_1$, only three distinct prefixes are
left: $PF_1 = \{a^6,\, e^6,\, b^6\}$ (we use the notation $c^n$ to
denote a length-$n$ string consisting of the same character $c$).
Then a new tree node $node_1$ is built adaptively upon $PF_1$ to hold
its prefixes as keys.  According to the first component of $tp_1$
(i.e. $root.aa$), $node_1$ is associated with the key $aa$ in the root
by a child pointer, which makes it to be the first child of root. Once
again, the suffixes left in $SF_1$ are grouped based on their lost
length-6 prefixes, which formed three new tuples:
$tp_4 = (node_1.a^6,\; SF_4=\{p_1^{-8}\})$,
$tp_5 = (node_1.e^6,\; SF_5=\{p_2^{-8},\, p_3^{-8},\, p_4^{-8}\})$ and
$tp_6 = (node_1.b^6,\; SF_6=\{p_5^{-8}\})$.

Note that, in order to mark the end of a pattern, once the last part
of the pattern has been removed and then stored in a node, that last
part is marked with an asterisk(*) in the corresponding node. As a
result of this, any path from the root to a key that marked with an
asterisk represents a complete pattern. It can be seen that, the
patterns are stored implicitly in the AMT and can be reconstructed
from the paths ended with an asterisk.

During the construction of AMT, we use a \emph{breadth-first} strategy
to process the sub-suffix-sets and create tree nodes (which means the
next sub-suffix-set to be processed is $SF_2$ rather than $SF_4$). For
the purpose of tracing the order in which the sub-suffix-sets are
processed, a \emph{first-in-first-out} queue is employed to maintain
the created tuples. The tuple at the beginning of the queue contains
the sub-suffix-set that will be processed next, while the newly
created tuples are inserted to the end of the queue in order. In our
example, the root node is built initially, then $tp_1$, $tp_2$ and
$tp_3$ are inserted to the queue. Next $SF_1$ of $tp_1$ is processed
and the newly created tuples $tp_4$, $tp_5$ and $tp_6$ are inserted to
the queue in order. Subsequently, $SF_2$, $SF_3$, $SF_4$, $SF_5$,
$SF_6$, $\dots$ are processed in sequence. Once there is no tuple left
in the queue, the whole AMT has been constructed.  It is worth
pointing out that, we can also use a depth-first strategy to build the
AMT branch by branch, but regardless of the strategies used, the final
built AMTs are the same.

\begin{algorithm}
  \caption{Constructing the AMT}\scriptsize
  \label{alg:amt}
  \begin{algorithmic}[1]
    \REQUIRE The pattern set $P$
    \ENSURE The corresponding AMT
    \STATE
    \STATE $Q \leftarrow$ Create an empty queue
    \STATE $push\_queue((NULL,\,P),\; Q)$
    \STATE
    \WHILE{$Q$ is not empty}
    \STATE $(parent\_node.key,\; SF) \leftarrow pop\_queue(Q)$
    \STATE $lss \leftarrow$ The length of the shortest pattern in $SF$
    \FOR{each $suf \in SF$}
    \IF{$|suf|=lss$}
    \STATE Mark $suf^{+lss}$ to be the pattern end
    \ENDIF
    \STATE Remove $suf^{+lss}$ of $suf$, and put $suf^{+lss}$ into $PF$
    \ENDFOR
    \STATE Discard the repetitions in $PF$, and let $ndp \leftarrow |PF|$
    \STATE According to $lss$ and $ndp$, adaptively create a
    $new\_node$ to hold the prefixes in
    $PF$
    \IF{$(parent\_node.key = NULL)$}
    \STATE $root \leftarrow new\_node$
    \ELSE
    \STATE Associate $new\_node$
    with $parent\_node.key$ by a child pointer
    \ENDIF
    \STATE $TP \leftarrow \{(new\_node.pf,\, SSF) \mid SSF \subseteq SF\; and
    \ \forall \ p,\,q \in SSF: p,\,q$ have the same length-$lss$ prefix
    $pf$ removed\}\
    \FOR{each $tp \in TP$}
    \STATE $push\_queue(Q,\,tp)$
    \ENDFOR
    \ENDWHILE
    \STATE
    \RETURN $root$.
  \end{algorithmic}
\end{algorithm}

The framework of constructing the AMT from a pattern set is
illustrated in Algorithm \ref{alg:amt}. Firstly, an empty queue $Q$ is
created in line 2 for maintaining the tuples. Then the tuple
$(NULL, P)$ is inserted to $Q$ in line 3 by function $push\_queue$
which always inserts an element to the end of the queue. Since $P$ is
the initial pattern set here, the first component of the tuple is
$NULL$, which means there is no prefix removed from $P$ yet. By this
way, the creation of the root can be combined with the creation of
other tree nodes in the following \textbf{while} loop.

If $Q$ is not empty, the \textbf{while} loop body from line 5 to line
25 constructs a tree node based on the first tuple in $Q$. The tuple
at the beginning of $Q$ is fetched by the $pop\_queue(Q)$ function in
line 6: the suffix set to be processed is assigned to $SF$, and the
corresponding prefix of $SF$ in the parent node is denoted by
$parent\_node.key$. Then the $lss$ of $SF$ is computed in line 7. In
the inner \textbf{for} loop from line 8 to line 13, the length-$lss$
prefix of each suffix in $SF$ is removed, and the prefixes are
collected to form the prefix set $PF$. If a prefix is the last part of
some pattern, it is marked as a pattern end. Line 14 discards the
repetitions in $PF$, after that, a new tree node is created adaptively
to hold the prefixes $PF$ in line 15. The \textbf{if} statement in
line 16 decides whether the newly created node is the root node or
not: if the prefix component of the current tuple is $NULL$, the new
node is the root node as stated before; otherwise, the new node is a
child node which will be then associated with the corresponding prefix
in its parent node. Next in line 21, the suffixes left in $SF$ are
grouped based on their lost prefixes, and each group and the
corresponding prefix forms a tuple. Finally, the created tuples are
inserted orderly to the end of $Q$ by the $push\_queue$ function. Once
there is no tuple left in $Q$, the whole AMT has been completely
constructed, and the root of AMT is returned.

\subsection{Verification}
\label{subsec:matching}

As stated in the filter module, once a suspicious position in the text
has been detected, the verification module checks that position for
real pattern occurrences. Specifically, for such a suspicious position
$i$, the engine searches a sequence of substrings of the text, where
the first substring starts at $i$, in the corresponding nodes of AMT
from the root. If any substring successfully matched a key that is
marked as a pattern end in some node, a pattern is declared to be
found at position $i$. However, if there is a mismatch or the
searching goes beyond the leaves of AMT, the engine immediately
forwards to the next position $i+1$ and invokes the filter again.

The pseudocode of the verification procedure is shown in Algorithm
\ref{alg:matching}. Suppose $i$ is a suspicious position of the text
detected by the filter module. The variable $m\_len$ is the total
length of the sub-strings of the text who have successfully matched
with keys in AMT nodes. The variable $node$ points to the current
matching node in the AMT. In addition, since all the keys in a node
have the same length, the notation $key\_len(node)$ is used to denote
the length of the keys in $node$.

\begin{algorithm}
  \caption{Verification}\scriptsize
  \label{alg:matching}
  \begin{algorithmic}[1]
    \REQUIRE ~~\\
    A suspicious position $i$ in the text $T$\\
    The AMT built from the pattern set $P$\\
    \ENSURE ~~\\
    The patterns (if any) start at $i$
    \STATE
    \STATE $m\_len \leftarrow 0$
    \STATE $node \leftarrow root$ of AMT
    \STATE
    \WHILE{$node \neq NULL$ and $\exists key \in node: key =
      T[i+m\_len, \, key\_len(node)]$}
    \STATE $m\_len \leftarrow m\_len + key\_len(node)$
    \IF{$key$ is marked as a pattern end}
    \STATE Report that the pattern $T[i,\,m\_len]$ is found at position $i$
    \ENDIF
    \STATE $node \leftarrow$ The child of $node$ corresponding to $node.key$
    \ENDWHILE
  \end{algorithmic}
\end{algorithm}

For the suspicious position $i$, the \textbf{while} loop from line 5
to line 11 checks whether there are patterns starting at $i$. The
checking starts form the root of AMT: if the corresponding sub-string
$T[i+m\_len, \, key\_len(node)]$ matches some $key$ in the current
$node$, the totally matched length $m\_len$ is increased by the length
of the key. At the same time, if the matched $key$ is also marked as a
pattern end, the verification module will report that a pattern,
i.e. $T[i,\,m\_len]$, is found at position $i$. Then the checking
transfers to the child of $node.key$ and makes that child node to be
the current node. Once there is a mismatch or the current node goes
beyond the leaves of AMT (i.e.  $node = NULL$), the checking
terminates, and the engine forwards to the next position $i+1$.

Note that, since the nodes in AMT have various types of inner
structures, the search of the target string in a node, must use the
search routine specified to the type of that node.

\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.75\textwidth]{./eps/match}
  \caption{The verification procedure for the position $i$ in the
    text.}
  \label{fig:matching}
\end{figure}

Next, an example is given in figure \ref{fig:matching} to illuminate
the verification for a suspicious position $i$ of a given text string
$T$, using the AMT built in section \ref{subsec:amt}. The verification
starts from the root of AMT: according to $key\_len(root)=2$, the same
length sub-string starting at position $i$ of $T$, which is
$T[i,\,2]=aa$, is taken to check whether it's one of the keys in
root. The result is: $aa \in root$, and $aa$ is also marked as a
pattern end. Therefore the pattern $T[i,\,2]=aa$ ($p_6$ in $P$) is
found at position $i$. Then the checking transfers to $node_1$, which
is the child node of $root.aa$.

At $node_1$, since $key\_len(node_1)=6$, the same length sub-string
$T[i+2,\,6]=e^6$, which follows $T[i,\,2]$, is trying to match with
some key in $node_1$. The result is that: $e^6 \in node_1$ but it is
not a pattern end. Then the checking simply transfers to $node_5$,
which is the child of $node_1.e^6$.

At $node_5$, since the next sub-string $T[i+8,\,5]=u^5 \in node_5$ and
$u^5$ is marked as a pattern end, another pattern
$T[i,\,13]=a^2e^6u^5$ ($p_4$ in $P$) is found at position $i$. Then
the checking goes to the child of $node_5.u^5$, i.e.  $node_{10}$.

At $node_{10}$, since $T[i+13,\,4]=sscc \notin node_{10}$, the
verification for position $i$ terminates. And next, the engine will
forward to position $i+1$ and invoke the filter module for that
position.

From above it can be seen that, each verification procedure actually
corresponds to a \emph{verification path} in the AMT, whose tree nodes
have been compared with the text in the former process. In particular,
the verification path of the given example is:
$root \rightarrow node_1 \rightarrow node_5 \rightarrow node_{10}$,
witch is indicated by the dashed arrows in Figure \ref{fig:matching}.

\subsection{Adaptive creation of the tree nodes}
\label{subsec:nodes}

As stated in section \ref{subsec:amt}, each node of AMT has a specific
index structure to hold the prefixes of a prefix set as the keys. The
type of the index structure is chosen adaptively according to the
features of the prefix set. Since all the keys in the prefix set have
the same length, the features of the prefix set can be mainly
characterized by two values: the \emph{length} and \emph{number} of
the keys in the prefix set. These two values have been respectively
symbolized by $lss$ and $ndp$ in Section \ref{subsec:amt}.

In this work, three different kinds of structures,
i.e. \emph{character map}, \emph{string array} and \emph{hash table}
can be adopted for prefix sets with different $lss$ and $ndp$
combinations. Next, the detailed descriptions of these structures will
be presented.

\subsubsection{Character map}

Given a prefix set, if the length of prefixes (keys) is equal to 1
($lss=1$), i.e. each key in the prefix set is just a single character,
the space-efficient \emph{character map} structure, which was proposed
by Leis \cite{Leis2013}, will be adopted as the tree node. There are
four sub-types of character maps with different capacities for
different number of keys ($ndp$), where $1 \leq ndp \leq 256$.

\begin{figure}[hp]
  \centering
  \includegraphics[height=3.5in, width=3.5in]{./eps/Maps}
  \caption{The four sub-types of character maps. The triangles
    represent the corresponding child nodes of the keys.}
  \label{fig:character map}
\end{figure}

Figure \ref{fig:character map} illustrates the four sub-types of
character maps which are named according to their maximum
capacity. Instead of using one array of (\emph{key, child pointer})
pairs, the key part and the child pointer part are stored in the
separate arrays, which is able to keep the node structure compact
while supporting efficient search.

\begin{itemize}
\item \textbf{Map 4} (for $1 \leq ndp \leq 4$): The smallest map type
  can store up to 4 keys. It uses an array of 4 entries for keys and
  another array of the same length for child pointers. The keys and
  child pointers are stored at corresponding positions in their
  arrays, and the keys are sorted according to their ACSII
  values. Once the target character is found in the key array, its
  child pointer can be located through the same position in the
  pointer array. Figure \ref{fig:character map} (a) shows a Map 4
  structure with three keys: $f$, $n$ and $p$, where the triangles
  with the keys inside represent the corresponding child nodes.

\item \textbf{Map 16} (for $5 \leq ndp \leq 16$): This map type is
  used for storing 5 to 16 keys. It has the similar structure as Map
  4, but both arrays have up to 16 entries. A target character can be
  retrieved efficiently by a binary search in the key array. Figure
  \ref{fig:character map} (b) shows a Map 16 structure with 14 keys.

\item \textbf{Map 48} (for $17 \leq ndp \leq 48$): As the number of
  keys increases, searching in the key array becomes
  expensive. Therefore, maps with more than 16 (but less than 49) keys
  do not store the keys explicitly. Instead, a 256-element \emph{index
    array} is used, which can be directly indexed by the ASCII value
  of the target character. This array stores only the array indexes
  (small inters in the range of $[0,47]$) of another pointer array
  that contains up to 48 child pointers. In this way, the storage
  space can be saved comparing with storing pointers directly, because
  each array index only requires one byte. Figure \ref{fig:character
    map} (c) shows a Map 48 structure, where the ASCII values of $a$,
  $b$ and $c$ are 97, 98 and 99 respectively.

\item \textbf{Map 256} (for $49 \leq ndp \leq 256$): The largest map
  type is simply an array of 256 child pointers with each pointer
  initialized to $NULL$. It is used for storing 49 to 256 keys.  In
  this kind of character map, the child node can be found directly
  through the array index which is the ASCII value of the target
  character. Different from other character maps, in Map 256, there is
  only one array, and it is not necessary to carry out the additional
  indirect access. Therefore, if most entries are not empty, this
  representation is also very space efficient. Figure
  \ref{fig:character map} (d) shows a Map 256 structure, where only
  the pointer array is adopted.
\end{itemize}

The pseudocode of searching in a node, whose type is character map, is
given in Algorithm \ref{alg:character map}.

\begin{algorithm}
  \caption{Searching in a node whose type is character map}\scriptsize
  \label{alg:character map}
  \begin{algorithmic}[1]
    \REQUIRE ~~\\
    A tree $node$ whose type is character map\\
    The target character $t\_ch$.
    \ENSURE ~~\\
    The child node of $node.t\_ch$ (possibly $NULL$).
    \STATE
    \STATE $count \leftarrow$ Number of keys in the character map
    \STATE
    \SWITCH{The type of the character map}
    \CASE{\textsf{Map 4}}
    \FOR{$i \leftarrow 0$ to $count-1$}
    \IF{$keys[i]=t\_ch$}
    \RETURN $child\_pointers[i]$
    \ENDIF
    \ENDFOR
    \STATE Report a mismatch.
    \ENDCASE
    \STATE
    \CASE{\textsf{Map 16}}
    \STATE $low \leftarrow 0, high \leftarrow count-1$
    \WHILE{$low \le high$}
    \STATE $mid \leftarrow \lfloor (low+high)/2 \rfloor$
    \IF{$t\_ch=keys[mid]$}
    \RETURN $child\_pointers[mid]$
    \ELSIF{$t\_ch<keys[mid]$}
    \STATE $high \leftarrow mid-1$
    \ELSE
    \STATE $low \leftarrow mid+1$
    \ENDIF
    \ENDWHILE
    \STATE Report a mismatch.
    \ENDCASE
    \STATE
    \CASE{\textsf{Map 48}}
    \IF{$index[t\_ch] \neq NULL$}
    \RETURN $child\_pointers[index[t\_ch]]$
    \ELSE
    \STATE Report a mismatch.
    \ENDIF
    \ENDCASE
    \STATE
    \CASE{\textsf{Map 256}}
    \RETURN $child\_pointers[t\_ch]$
    \ENDCASE
    \ENDPWITCH
  \end{algorithmic}
\end{algorithm}

\subsubsection{String array}
\label{sec:string array}

For the prefix set whose length of keys is greater than 1 ($lss > 1$)
and the number of keys is not greater than 100 ($ndp \leq 100$), the
\emph{string array} structure is adopted as a tree node. Similar to
the Map 4 and Map 16 structures, the keys are stored in
lexicographical order in a separate key array with $ndp \times lss$
bytes, in which each key takes $lss$ bytes. The child pointers are
stored at the corresponding positions in another pointer array. Figure
\ref{fig:string array} illustrates a string array structure holding
three keys: $aaa$, $bbb$ and $ccc$.

\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.45\textwidth]{./eps/string_array}
  \caption{The string array structure with three keys: $aaa,\, bbb,\,
    ccc$.}
  \label{fig:string array}
\end{figure}

For efficiency, if the number of keys is less than 5, a naive linear
search is used to search the target string in the key array;
otherwise, a more efficient (but complicated) binary search algorithm
is adopted.  The pseudocode of the searching in a node whose type is
string array is depicted in Algorithm \ref{alg:string array}, where we
use the notation $A \prec B$ to denote that string $A$ is
\emph{lexicographically smaller} than string $B$. For the string array
whose number of keys is less than 5, the target string is compared
sequentially with the keys in the key array. If the target string
matched some key, return the child pointer at the corresponding
position in the pointer array; otherwise, report a mismatch. On the
other hand, if the number of keys is greater than 4, the target string
is searched by a binary search routine from the middle element of the
key array.

\begin{algorithm}
  \caption{Searching in a node whose type is string array}\scriptsize
  \label{alg:string array}
  \begin{algorithmic}[1]
    \REQUIRE ~~\\
    A tree $node$ whose type is string array; \\
    The target string $t\_str$.\\
    \ENSURE ~~\\
    The child node of $node.t\_str$ (possibly $NULL$).\\
    \STATE
    \STATE $count \leftarrow$ The number of keys in $node$
    \STATE
    \IF{$count < 5$}
    \STATE $i \leftarrow 0$
    \WHILE{$i < count$ and $keys[i] \prec t\_str$}
    \STATE $i \leftarrow i+1$
    \ENDWHILE
    \IF{$i < count$ and $keys[i]=t\_str$}
    \RETURN $child\_pointers[i]$
    \ELSE
    \STATE Report a mismatch
    \ENDIF
    \ELSE
    \STATE $low \leftarrow 0$, $high \leftarrow count-1$
    \WHILE{$low \leq high$}
    \STATE $mid \leftarrow \lfloor (low+high)/2 \rfloor$
    \IF{$t\_str = keys[mid]$}
    \RETURN $child\_pointers[mid]$
    \ELSIF{$t\_str < keys[mid]$}
    \STATE $high \leftarrow mid - 1$
    \ELSE
    \STATE $low \leftarrow mid + 1$
    \ENDIF
    \ENDWHILE
    \STATE Report a mismatch.
    \ENDIF
  \end{algorithmic}
\end{algorithm}

\subsubsection{Hash table}
\label{sec:hash table}

For the prefix set whose number of keys is greater than 100 ($lss > 1$
and $ndp > 100$), searching in the string array structure becomes
inefficient even using a binary search. In this case, a more fast (but
heavy) data structure --- \emph{hash table} is adopted to deal with
large number of keys. It is worth noting that, the construction of the
hash table is directly based on the suffix set from which the prefix
set is formed rather than the prefix set which is the basis of
building other node structures. In our implementation, a hash table is
an array of child pointers with each pointer initialized to $NULL$,
and we also utilize a \emph{string hash function} which transforms a
string to a positive integer.

% Indeed, when computing $ndp$, we
% merely count rather than really constructing the prefix set and then
% computing its $ndp$

In particular, given a suffix set $SF$, the size of the hash table
(denoted by $table\_size$) is determined by $ndp$ and a given load
factor $lf$ (ratio of $ndp$ to $table\_size$), i.e.
$table\_size = \lceil ndp\,/\,lf \rceil$. For example, with the $ndp$
of 1000 and a load factor of $70\%$, the $table\_size$ is
$\lceil 1000/0.7 \rceil = 1429$. Note that, the $ndp$ here is defined
to be the number of distinct length-$lss$ prefixes of the suffixes in
$SF$, which is equal to the size of the prefix set derived from $SF$
(after removing the repetitions).

Algorithm \ref{alg:hash} shows the pseudocode of building a hash table
based on $SF$. $SF$ is firstly partitioned into small suffix sets by
hashing: for each suffix $suf \in SF$, its prefix $suf^{+lss}$ is
hashed to an integer $i$ between 0 and $table\_size-1$ by the string
hash function, then $suf$ is associated with the $i$-th slot of the
hash table. After that, the suffixes whose length-$lss$ prefixes
hashing to the same value are associated with the same slot of the
hash table. Then, to address the hash collisions, for each slot that
is not $NULL$, the associated suffixes form a small suffix set $SSF$,
and a new tree node is created adaptively based on the prefix set
derived from $SSF$, as we have shown in the \textbf{while} loop of
Algorithm \ref{alg:amt}. The newly created tree node is then
associated with the corresponding slot of the hash table. As a result,
many tree nodes with various types can be associated within one hash
table.

% The suffixes whose length-$lss$ prefixes hashed to the same
% value $i$ are collected together associated with the $i$-th slot and
% forms a small suffix set. (Note that, we do not remove the prefixes of
% the suffixes when partitioning $SF$.) Then, to address the collisions,
% for each non-$NULL$ slot of the hash table, a tree node is built based
% on the small suffix set associated with that slot as stated
% before. Hence, many tree nodes with various structures can be
% incorporated in one hash table.

\begin{algorithm}
  \caption{Building a hash table}\scriptsize
  \label{alg:hash}
  \begin{algorithmic}[1]
    \REQUIRE ~~\\
    The suffix set $SF$ and the load factor $lf$. \\
    \ENSURE ~~\\
    The hash table.\\
    \STATE
    \STATE $ndp \leftarrow$ The number of distinct length-$lss$
    prefixes of suffixes in $SF$
    \STATE $table\_size \leftarrow \lceil ndp\,/\,lf \rceil$
    \STATE $hash\_table \leftarrow$ Create an array  of
    $table\_size$ pointers with each pointer initialized to $NULL$
    \STATE
    \FOR{each $suf \in SF$}
    \STATE $i \leftarrow Hash(suf^{+lss})$
    \STATE Associate $suf$ with $hash\_table[i]$
    \ENDFOR
    \STATE
    \FOR{$i \leftarrow 0$ to $table\_size - 1$}
    \IF{$hash\_table[i]\, \neq\, NULL$}
    \STATE $SSF \leftarrow \{suf\,|\,suf\in SF\; and\; Hash(suf^{+lss})=i\}$
    \STATE $new\_node \leftarrow$ Adaptively create a tree node from
    $SSF$, as shown in the \textbf{while} loop of Algorithm \ref{alg:amt}
    \STATE Associate $new\_node$ with $hash\_table[i]$
    \ENDIF
    \ENDFOR
    \STATE
    \RETURN $hash\_table$
  \end{algorithmic}
\end{algorithm}

Given a target string of length $lss$, the hash value of that string
is computed and used as the index of the hash table. If the
corresponding slot is $NULL$, which means there is a mismatch, the
verification terminates and the engine moves to the next position;
otherwise, the process goes to the tree node associated with that slot
and compares the node with the corresponding substring in the text.

The string hash function adopted can have a significant effect on the
performance of verifying. In our implementation, for each matching
task, the hash function is selected randomly from a \emph{hash
  function family} $H$. The hash function family we adopted is the
fast \emph{shift-add-xor} hash family proposed by Ramakrishna
\cite{Ramakrishna1997}, which is claimed to be both uniform and
universal.

\subsection{Further Improvements}
\label{sec:further improments}

The verification module presented above is pretty efficient, but there
are still some techniques that can further improve the time and space
efficiency of this module.

\subsubsection{Dividing the AMT}
\label{sec:divide amt}

The verification module we have just built is based on a big single
AMT which is constructed from the whole pattern set. In fact, we can
divide the big AMT into small ones and enable each verifying procedure
to refer to only one small AMT, by which the time for verifying can be
reduced. The division is based on the following observations.

As stated in Section \ref{sec:filter}, the filtering is based on the
last $k$-symbol block of the substring in the sliding window. In
particular, for a position $i$ of the text $T$, if some pattern $p_j$
starts at $i$, there must be
$Hash[T[i+lss-k,k]] = Hash[p_j[lss-k+1,k]]$ (we call
$Hash[p_j[lss-k+1,k]]$ the \emph{signature} of $p_j$). This means that
we can group the patterns according to their signatures: patterns with
the same signature are grouped and form a sub-pattern-set, and for
each sub-pattern-set we build a (small) AMT. In addition, an index
table is adopted to allow the corresponding AMT be retrieved in $O(1)$
time by its signature. Now, the verification module consists of an
index table as well as several small AMTs.

In the filtering phase, once a suspicious position $i$ in $T$ is
detected, the value $Hash[T[i+lss-k,k]]$ will be used as the index of
the index table to find the corresponding small AMT for verifying. As
the scale of the AMT shrunk, the time for each verifying is reduced
accordingly.

\subsubsection{Node Merging}
\label{sec:node merge}

Similar to the ``path compression'' technique that is widely used in
many trie-like structures, the nodes in a path of AMT, each of which
has only one key, can be merged into one node. This technique can
improve both time and space efficiency of AMT. To illustrate its
benefits, consider the pattern set
$P=\{p_1=e^6u^5sscc,\; p_2=e^6u^5,\; p_3=e^6\}$ and the corresponding
AMT constructed by Algorithm 1. As shown in Figure \ref{fig:merge}, the
built AMT has three nodes, and each contains only one key. In a
typical x86\_64 architecture, this implementation of AMT takes roughly
55 bytes of memory: 15 bytes for the three keys, 16 bytes for the two
child pointers and 24 bytes for the three pointers to the matching
functions. However, if we combine the three keys into one string,
these three nodes can be merged into one node, as a result, the child
pointers are no longer needed and the number of function pointers is
reduced from three to one. The merged node is another type of tree
node in addition to the node types we have introduced in Section
\ref{subsec:nodes}.

\begin{figure}[htbp]
  \centering
  \includegraphics[height=2in, width=2in]{./eps/node_merge}
  \caption{Node merging.}
  \label{fig:merge}
\end{figure}

For the merged node, the only problem is how to distinguish among the
keys after they are concatenated. To address this issue, a
\emph{length array} is used to store the length of each key, where
each array element takes only one byte (using the \emph{unsigned char}
type). Through the length of each key, we can delimit each original
key from others, and thus, searching the target string in the merged
node can be trivial.

In general, after node merging, the AMT contains just one merged node
which only takes 26 bytes (15 bytes for the keys, 3 bytes for the
length array, and 8 bytes for the matching function pointer), thus, 29
bytes are saved comparing with the original AMT. In addition, only one
random access to the memory is needed to get the merged node (the
original AMT needs three random memory accesses to get all its nodes),
thus, the cache miss is reduced and the performance is improved.

\subsubsection{Node splitting}
\label{sec:node split}

In order to further improve the searching efficiency of the tree
nodes, once the keys in a node share a common prefix, that prefix
should best be removed from the keys and form a new single key node
. (Note that, since the prefix is either a character or a string, the
type of the new created node is either Map 4 or String Array.)  This
action can reduce a lot of redundant comparisons while searching a
target key in a node. For instance -- as shown in Figure \ref{fig:split},
consider searching a target key $a^5u$ in a node of String Array type,
where the four keys of the node share a common prefix $a^5$. According
to Algorithm \ref{alg:string array}, the target key have to be
compared sequentially with the four keys, which will totally takes
$6 \times 4 = 24$ byte comparisons to determine the inexistence of
target key. However, if the common prefix $a^5$ is removed out of the
keys and form another new node as shown in Figure \ref{fig:split}, for
the same target key, only 9 byte comparisons are needed for verifying
(5 comparisons for $node_1$ and another 4 comparisons for
$node_2$). Obviously, comparing with only one node, the two nodes
version can reduce 15 comparisons.

\begin{figure}[htbp]
  \centering
  \includegraphics[height=2.5in, width=2in]{./eps/node_split}
  \caption{Node splitting.}
  \label{fig:split}
\end{figure}

\section{Experimental Results}
\label{sec:experiments}

In this section, we evaluate the performance of the proposed fast
engine for multi-pattern matching (\textsf{FEMPM} for short), and
compare it with other five algorithms: the \textsf{AC} and \textsf{WM}
algorithms, which are two classic ``baseline'' algorithms for MPM; the
\textsf{Split}, \textsf{MASM} and \textsf{Pre-filter+AC} algorithms,
which are three state-of-the-art MPM algorithms with good efficiency
in practice. All these algorithms are evaluated in two aspects: the
robustness of these algorithms under pattern sets with various $lsp$s
and the scalability of the algorithms under pattern sets with various
numbers of patterns.

\subsection{Simulation Settings}

The experiments are conducted on a PC with an Intel core i7 2.93GHz
CPU, 8GB of RAM and 1TB Disk Driver; the operating system is Windows 7
Professional (64-bits). All the testing algorithms are implemented in
C/C++ within the IDE of \emph{Code::Blocks}. The test data used in the
experiments is the real-world English text from the Pizza\;\&\;Chili
corpus \cite{pizzachil}. The patterns are extracted randomly from the
text to form the pattern sets.

Next, the parameter configuration of \textsf{FEMPM} in the experiment
is given as follows: for the String Arrays, if the number of strings
exceeds \textbf{4}, the binary search is used to search the target
string in the key array; otherwise, the linear search is adopted (for
String Arrays containing no more than 4 strings, the simple linear
search is faster than binary search due to the complexity of binary
search). For the Hash Tables, the load factor is set to be
\textbf{0.5}, which is a good balance between time and space; the seed
of the hash function is generated randomly in the range of 1~50; the
shift values L and R are set to \textbf{2} and \textbf{6} respectively
(the parameters about the hash function are chosen from
\cite{Ramakrishna1997} which are claimed to be very efficient). Once
the $ndp$ is larger than \textbf{100}, the Hash Table becomes more
efficient than the String Array, thus the Hash Table is built once
$ndp>100$. These choices of parameters have been extensively tested on
a wide range of pattern sets, and are considered to be very efficient
in practice.

\subsection{Robustness Evaluation}

As we've mentioned before, many MPM algorithms are sensitive to the
length of the shortest pattern ($lsp$) in the pattern set. So first,
it is necessary to measure the robustness of \textsf{FEMPM} and other
investigated algorithms under pattern sets with various $lsp$s. In
this evaluation, there are totally 9 pattern sets for testing with
$lsp$s ranging from 2 to 10, and each pattern set has a fixed number
of $10^5$ patterns. The size of text string is fixed to 200 MB. The
characteristics of the testing pattern sets are shown in Table
\ref{tab:lsps}, which includes: the length of the shortest pattern
(LSP), the length of the longest pattern (LLP), the average length of
the patterns (ALP), the standard deviation of the pattern lengths
(SD), the total length of the patterns (TLP) and the number of
patterns in the pattern set (Count).

\begin{figure}[htbp]
  \centering
  \includegraphics[height=2in, width=2.5in]{./eps/lsp}
  \caption{Matching times for various lsp.}
  \label{fig:lsp}
\end{figure}

\begin{table}
  \centering
  \caption{The characteristics of pattern sets with various LSPs.}
  \scriptsize
  \label{tab:lsps}
  \begin{tabular}{p{20pt}p{10pt}p{10pt}p{10pt}p{10pt}p{30pt}p{20pt}}
    \hline
    Pattern Set & LSP  & LLP  & ALP & SD & TLP & Count\\
    \hline
    $P_1$ & 2 & 50 & 25.9 & 14.1 & 2,599,020 & $10^5$\\
    $P_2$ & 3 & 50 & 26.5 & 13.9 & 2,646,762 & $10^5$\\
    $P_3$ & 4 & 50 & 27.0 & 13.6 & 2,704,091 & $10^5$\\
    $P_4$ & 5 & 50 & 27.6 & 13.2 & 2,745,545 & $10^5$\\
    $P_5$ & 6 & 50 & 27.9 & 13.0 & 2,793,178 & $10^5$\\
    $P_6$ & 7 & 50 & 28.4 & 12.7 & 2,843,580 & $10^5$\\
    $P_7$ & 8 & 50 & 29.0 & 12.4 & 2,903,343 & $10^5$\\
    $P_8$ & 9 & 50 & 29.5 & 12.1 & 2,948,372 & $10^5$\\
    $P_9$ &10 & 50 & 30.0 & 11.8 & 2,998,992 & $10^5$\\
    \hline
  \end{tabular}
\end{table}

Figure \ref{fig:lsp} presents the mean matching time (in seconds) used
by these algorithms based on 5 independent runs. In general, the
\textsf{AC} algorithm is not affected much by the changes in
$lsp$. However, the DFA built by this algorithm consumes a huge amount
of memory. What's worse is that, the frequent and unpredictable
state/failure transitions will lead to a lot of cache misses during
matching, and these result in the low performance of \textsf{AC}. On
the other side, the \textsf{Split} algorithm, which is also based on
DFA, is not effected much by $lsp$, and since it extracts a subset of
unique patterns to build the DFA and uses a bit-split technique, the
memory and time consumption are greatly reduced. It can be seen from
Figure \ref{fig:lsp} that the \textsf{WM} algorithm is affected much
by the changes in $lsp$: it outperforms \textsf{AC} on the pattern
sets whose $lsp$ is larger than 2, but for the pattern sets with small
$lsp$, the performance of \textsf{WM} degrades dramatically. This is
due to the ``skip strategy'' of \textsf{WM} algorithm that does not
work well for the pattern sets with small $lsp$s, which means most of
the positions in the text string need to be checked by the HASH table
of \textsf{WM}, and this takes a lot of time. The
\textsf{Pre-filter+AC} algorithm outperforms all other algorithms
except for \textsf{FEMPM} for pattern sets whose $lsp$ is larger than
5. Similarly, this algorithm also uses a skip strategy: once a
position of the text string is filtered out by the Pre-filter, at most
the following $lsp-1$ positions can be skipped over. Consequently, for
pattern sets with small $lsp$, this algorithm can not skip much in the
text string, which will seriously affect its performance. In contrast,
the \textsf{MASM} algorithm performs well for small $lsp$
($lsp \leq 4$). This algorithm first compacts the pattern set by a
prefix tree: if a pattern $A$ is a prefix of another pattern $B$, $A$
can be merged into $B$. Then the algorithm builds a binary search tree
for the compacted pattern set based on lexicographical
order. Comparing with the long patterns, the short ones are more
likely to be prefixes of others, and thus, more likely to be
eliminated by compacting. Thus, pattern sets with small $lsp$ are well
compressed, which will improve both time and space efficiency of
\textsf{MASM} in the matching phase.

From Figure \ref{fig:lsp} we can also see that, \textsf{FEMPM}
outperforms all other investigated algorithms on every
$lsp$. Comparing with \textsf{AC}, the matching time is reduced by
$85\%$ in average for all $lsp$s; even comparing with the fast
\textsf{Pre-filter+AC} algorithm, the matching time is reduced by
$10\% \sim 20\%$ for medium $lsp$s ($lsp \geq 4$) and by $70\%$ for
small $lsp$s ($lsp < 4$). Although the filter part of \textsf{FEMPM},
which likes that of \textsf{Pre-filter+AC}, is not suitable for small
$lsp$, the verification part, i.e. AMT, is more efficient than that of
the \textsf{Pre-filter+AC}. This is mainly due to the adaptivity of
the nodes of AMT, which enables the tree nodes as efficient and
compact as possible. Therefore, no matter what the $lsp$s of the
pattern sets are, the corresponding AMT is able to keep high search
efficiency as well as low memory requirement in the verification
phase. Owing to this, the \textsf{FEMPM} has the best robustness for
all of the testing pattern sets. The statistical data about node types
of corresponding AMTs are shown in table \ref{tab:node types}. The Map
1 and Single String Array types in the table are actually the Map 4
and String Array types which contain only one element,
respectively. Obviously, as the $lsp$ of the pattern set changes, the
number of nodes of the each type in the AMT changes accordingly, which
reflects the adaptivity of AMT.

Moreover, the performance of \textsf{FEMPM} is very stable and
reliable. In particular, once the length of the text and number of
patterns are fixed, the matching time changes little for various
$lsp$s. As shown in Figure \ref{fig:lsp}, the matching time of
\textsf{FEMPM} keeps almost 10 seconds unchanged for a text of about
200MB and a pattern set with about $10^5$ patterns; for $lsp$s ranging
from $2$ to $10$, the change in the matching time is less than 2
seconds. In fact, as the $lsp$ increases, the average length of the
paths from the root to the leaves in AMT will increase
slightly. Accordingly, in the matching phase, the matching paths for
some text positions may get a little bit longer, which might slightly
affect the matching speed.

\begin{table}[!htp]
  \centering
  \caption{The statistics about the AMTs under various lsps}
  \scriptsize
  \label{tab:node types}
  \begin{tabular}{p{10pt}p{15pt}p{15pt}p{15pt}p{15pt}p{15pt}p{25pt}p{25pt}p{20pt}p{25pt}}
 \hline
 Lsp &
 Map 1 &
 Map 4 &
 Map 16 &
 Map 48 &
 Map 256 &
 Single String Array &
 String Array   &
 Hash Table &
 Total\\
 \hline
 2  & 2,464 & 2,096 & 663 & 133 & 0 & 63,274 &  21,627 & 10 & 90,267\\
 3  & 2,379 & 1,636 & 385 & 90  & 0 & 64,701 &  22,551 & 10 & 91,752\\
 4  & 2,273 & 1,329 & 201 & 57  & 0 & 66,162 &  22,736 &  1 & 92,759\\
 5  & 2,045 & 1,104 & 112 & 45  & 0 & 68,152 &  22,048 &  1 & 93,507\\
 6  & 1,758 &   899 &  52 & 34  & 0 & 70,179 &  20,818 &  1 & 93,741\\
 7  & 1,560 &   807 &  29 & 33  & 0 & 72,284 &  19,280 &  1 & 93,994\\
 8  & 1,515 &   803 &  20 & 31  & 0 & 73,447 &  18,274 &  1 & 94,091\\
 9  & 1,441 &   788 &  24 & 30  & 0 & 74,584 &  17,120 &  1 & 93,988\\
10  & 1,361 &   794 &  24 & 29  & 0 & 75,094 &  16,427 &  1 & 93,730\\
\hline
  \end{tabular}
\end{table}

\subsection{Scalability Evaluation}

In this section, we evaluate the scalability of the investigated
algorithms by testing them under pattern sets with various numbers of
patterns. As before, the size of the text string is fixed to 200MB,
and there are two groups of pattern sets for testing --- the small
group and the large group. The small group contains 9 pattern sets,
and the sizes (number of patterns) of the pattern sets increase from
$1 \times 10^5$ to $9 \times 10^5$ with step size of $10^5$, while the
large group contains 10 pattern sets, and the sizes increase from
$10^6$ to $10^7$ with step size of $10^6$. The range of the pattern
length of each pattern set is fixed to $5 \sim 50$.

\begin{figure}[htbp]
  \centering
  \includegraphics[height=2in, width=2.5in]{./eps/small_group}
  \caption{Matching times for the small group.}
  \label{fig:small_group}
\end{figure}

For the small group, the mean matching times of the investigated
algorithms under 5 independent runs are shown in Figure
\ref{fig:small_group}. The experimental results indicate that, the
performance of the \textsf{AC} algorithm is not good but relative
stable with the number of patterns growing. Also it can not deal with
pattern sets whose size is larger than $7 \times 10^5$ in this testing
, due to out of memory. Although the \textsf{WM} algorithm performs
better than \textsf{AC}, its performance is not stable. For instance,
the matching time only rises about $1s$ as the number of patterns
increases from $2 \times 10^5$ to $3 \times 10^5$, but when the number
of patterns increases from $8 \times 10^5$ to $9 \times 10^5$, the
matching time rises up to $9s$. Thus, we can not estimate the matching
time of \textsf{WM} based on the size of the pattern set. The
\textsf{Pre-filter+AC} performs well as the number of patterns grows,
but its matching time is also unpredictable, e.g. the matching times
are nearly the same as the number of patterns increases from
$4 \times 10^5$ to $5 \times 10^5$. The reason is that: the effect of
the filters used by these two algorithms highly depends on the
contents of the text and patterns themselves. For some pattern sets
whose patterns occurs frequently in the text, the effect of the
filters degrades. On the other hand, as the number of patterns grows,
the matching time of \textsf{MASM} grows more slowly than that of the
\textsf{Pre-filter+AC}, which is due to that the matching time of
\textsf{MASM} mainly relies on the depth of the pipelinded binary
search tree and that depth grows very slow as the number of patterns
increases. The \textsf{Split} algorithm performs very well for the
small group of patterns, it is only $10\% \sim 20\%$ slower than our
\textsf{FEMPM}.

\begin{table}[!htp]
  \scriptsize
  \caption{The statistics about the node type in AMTs built for the small group}
  \label{tab:small}
  \begin{tabular}{p{25pt}p{25pt}p{25pt}p{25pt}p{15pt}p{15pt}p{25pt}p{25pt}p{20pt}p{25pt}}
 \hline
 Size &
 Map 1 &
 Map 4 &
 Map 16 &
 Map 48 &
 Map 256 &
 Single String Array &
 String Array   &
 Hash Table &
 Total\\
\hline
$1 \times 10^5$  &  2,000 &  1,006 &    72 &  18 & 0 &  68,268 &  21,971 & 103 &  93,438 \\
$2 \times 10^5$ &  4,628 &  2,808 &   307 &  42 & 0 & 130,440 &  46,514 & 244 & 184,983 \\
$3 \times 10^5$ &  7,414 &  4,858 &   574 &  84 & 0 & 190,653 &  71,757 & 403 & 275,746 \\
$4 \times 10^5$ & 10,555 &  7,282 &   946 & 139 & 0 & 247,656 &  97,967 & 544 & 365,089 \\
$5 \times 10^5$ & 13,703 & 10,651 & 1,503 & 238 & 4 & 302,350 & 124,528 &  17 & 452,994 \\
$6 \times 10^5$ & 17,335 & 13,536 & 1,954 & 298 & 0 & 356,644 & 151,435 &  31 & 531,233 \\
$7 \times 10^5$ & 21,080 & 16,872 & 2,405 & 388 & 4 & 408,165 & 179,621 &  35 & 628,570 \\
$8 \times 10^5$ & 24,357 & 20,094 & 2,977 & 436 & 3 & 457,674 & 208,426 &  37 & 714,004 \\
$9 \times 10^5$ & 28,356 & 23,924 & 3,441 & 513 & 9 & 504,779 & 238,632 &  42 & 799,696 \\
\hline
  \end{tabular}
\end{table}

Among all these algorithms, the \textsf{FEMPM} is the most efficient
as well as stable one. The high performance of \textsf{FEMPM} is
mainly due to the \emph{root} node of AMT, which is almost always the
Hash Table as the number of patterns grows. As mentioned before, the
root plays the role of another ``filter'', by which a large number of
positions of the text that can not be filtered by the filter part may
have a chance to be filtered out by the root. Moreover, the increasing
in the number of patterns mainly leads to the growing of the width
rather than the depth of AMT. Therefore, for each position of the
text, the checking time changes slightly. From the experimental
results, we can even give a rough estimate on the matching time based
on the size of pattern set as follows: given the text of 200 MB, an
increase of $10^5$ in the number of patterns will lead to roughly $1$
more second taken in the matching time. The statistics about the node
types of the corresponding AMTs for the small group are shown in Table
\ref{tab:small}. We can see that an increase of $10^5$ in the number
of patterns will yield about $9 \times 10^4$ tree nodes.

\begin{figure}[htbp]
  \centering
  \includegraphics[height=2in, width=2.5in]{./eps/large_group}
  \caption{Matching times for the large group.}
  \label{fig:large_group}
\end{figure}

For the large group, the mean matching times of the investigated
algorithms are shown in Figure \ref{fig:large_group}. It can be seen
that, as the number of patterns grows from $10^6$ to $10^7$, the
increases in the matching time of the investigated algorithms are:
$410\%$ (\textsf{WM}), $374\%$ (\textsf{Pre-filter+AC}), $238\%$
(\textsf{MASM}), $191\%$ (\textsf{Split}), $112\%$ (\textsf{FEMPM})
respectively. For every increase of $10^6$ in the number of patterns,
the mean increment in the matching time of the investigated algorithms
are: $23.7s$ (\textsf{WM}), $17.0s$ (\textsf{Pre-filter+AC}), $12.3s$
(\textsf{MASM}), $8.8s$(\textsf{Split}), $3.5s$ (\textsf{FEMPM})
respectively. This indicates that, \textsf{Pre-filter+AC} and
\textsf{WM} perform not well for large pattern sets. This is because,
as the number of patterns grows, the chances for every position of the
text to successfully match a pattern increases, which will reduce the
effect of the filters of these algorithms and results in performance
degradation. The \textsf{MASM} algorithm outperforms
\textsf{Pre-filter+AC} for pattern sets whose size are larger than
$3 \times 10^6$, which is due to its pattern set compression strategy
and the good scalability of its pipelinded binary search tree. The
\textsf{Split} algorithm has a better scalability than others (except
\textsf{FEMPM}), this is because it only extracts a subset from the
whole pattern set, in which a pattern is not a suffix of any other
patterns, to build the DFA.

\begin{table}[!htp]
  \caption{The statistics about the node type in the AMTs built for
    the large group}
  \scriptsize
  \label{tab:large_group}
  \begin{tabular}{p{25pt}p{25pt}p{25pt}p{25pt}p{20pt}p{10pt}p{30pt}p{30pt}p{20pt}p{25pt}}
 \hline
 Size &
 Map 1 &
 Map 4 &
 Map 16 &
 Map 48 &
 Map 256 &
 Single String Array &
 String Array   &
 Hash Table &
 Total\\
\hline
$1 \times 10^6$ &  38,935 &   27,428  &   6,468 &     851   &    4 &    588,343  &    240,886 &  1,027 &    903,942  \\
$2 \times 10^6$ &  85,900 &   68,973  &  17,084 &   2,141   &   20 &  1,091,527  &    496,001 &  2,015 &  1,763,661  \\
$3 \times 10^6$ & 136,054 &  117,527  &  28,092 &   3,410   &   17 &  1,547,659  &    764,359 &  2,749 &  2,599,867  \\
$4 \times 10^6$ & 189,926 &  171,155  &  40,328 &   4,716   &   23 &  1,966,326  &  1,045,558 &  3,450 &  3,421,482  \\
$5 \times 10^6$ & 247,707 &  230,163  &  83,043 &   5,989   &  38 &  2,354,308   &  1,332,896 &  3,885 &  4,228,029  \\
$6 \times 10^6$ & 304,863 &  293,591  &  65,475 &   7,169   &   49 &  2,706,094  &  1,625,261 &  4,379 &  5,006,881  \\
$7 \times 10^6$ & 366,861 &  361,695  &  77,899 &   8,434   &   53 &  3,030,673  &  1,925,027 &  4,733 &  5,775,380  \\
$8 \times 10^6$ & 429,121 &  433,708  &  90,280 &   9,660   &   62 &  3,336,765  &  2,226,879 &  5,070 &  6,531,545  \\
$9 \times 10^6$ & 494,278 &  509,992  & 102,402 &  10,833   &   72 &  3,623,407  &  2,537,413 &  5,224 &  7,283,621  \\
$1 \times 10^7$ & 558,241 &  591,455  & 115,012 &  11,922   &   79 &  3,986,683  &  2,847,277 &  5,505 &  8,026,174  \\
\hline
\end{tabular}
\end{table}

On the other hand, among all the algorithms, \textsf{FEMPM} is the
most efficient algorithm for large pattern sets. An increase of $10^6$
in the number of patterns, only requires 2.8 more seconds to match,
this is mainly due to the good scalability of the AMT which enables
the matching time grows very slow as the number of patterns
explodes. The statistical data about the AMTs for the large group are
shown in Table \ref{tab:large_group}. Every $10^6$ increase in the
number of patterns will increase only about $8 \times 10^5$ tree
nodes.

\section{Conclusion and Future Works}
\label{sec:conclusion}

In this paper, we propose a Fast Engine for Multi-Pattern Matching
(\textsf{FEMPM}) which includes a filter part and a verification
part. The filter part is used to quickly filter out the positions that
are impossible to match any pattern, while the verification part is
used to verify the potential positions. The verification part used a
data structure called Adaptive Matching Tree (AMT) which can be
constructed to fit the feature of the given pattern set. Owing to the
scalability and compactness of AMT, \textsf{FEMPM} offers a good
support for large-scale pattern sets. Moreover it can improve the
robustness of \textsf{FEMPM} for pattern sets with various $lsp$s. In
the future, we will try to design more efficient inner structures of
the tree nodes.

\section{Acknowledgments}

This work is supported by National Natural Science Foundation of China
(No.61472297) and the Fundamental Research Funds for the Central
Universities (BDZ021430).

\begin{thebibliography}{99}

\bibitem{Yan2016} Yan, Ruoyu. ``Combining Renyi Entropy and EWMA to
  Detect Common Attacks in Network'' {\it International Journal of
    Pattern Recognition and Artificial Intelligence}, 30(10),
  pp.1650021, 2016.

\bibitem{Xue2015} Xue, Xingsi and Wang, Yuping, ``Optimizing ontology
  alignments through a Memetic Algorithm using both MatchFmeasure and
  Unanimous Improvement Ratio'', {\it ARTIFICIAL INTELLIGENCE}, 223,
  pp.65-81, 2015.

\bibitem{Xue2016} Xue, Xingsi and Wang, Yuping, ``Using Memetic
  Algorithm for Instance Coreference Resolution'', {\it IEEE
    TRANSACTIONS ON KNOWLEDGE AND DATA ENGINEERING}, 223, pp.65-81,
  2015.

\bibitem{Kim2015} Kim, HyunJin and Choi, Kang-Il and Choi, Sang-Il,
  ``A Memory-Efficient Deterministic Finite Automaton-Based Bit-Split
  String Matching Scheme Using Pattern Uniqueness in Deep Packet
  Inspection'', {\it PLOS ONE}, 10(5), 2015.

\bibitem{Xiao2016} Guoqing Xiao, Kenli Li, Xu Zhou, Keqin
  Li. ``Queueing Analysis of Continuous Queries for Uncertain Data
  Streams Over Sliding Windows'' {\it International Journal of Pattern
    Recognition and Artificial Intelligence}, 30(9), pp.1660001, 2016.

\bibitem{Wu1994} Sun Wu and Udi Manber, ``A fast algorithm for
  multi-pattern searching'', {\it Technical Report-94-17}, 1994.

\bibitem{Lee2013} Tsern-Huei Lee and Nai-Lun Huang, ``A
  pattern-matching scheme with high throughput performance and low
  memory requirement'', {\it IEEE/ACM Transactions on Networking},
  21(4), pp.1104--1116, 2013.

\bibitem{Aho1975} Alfred V. Aho and Margaret J. Corasick, ``Efficient
  string matching:An aid to bibliographic search'', {\it Commun. ACM}
  18(6), pp.333-340, 1975.

\bibitem{Knuth1977} Donald E. Knuth, Jr. James H. Morris, and Vaughan
  R. Pratt, ``Fast pattern matching in strings'', {\it SIAM Journal on
    Computing}, 6(2), pp.323–350, 1977.

\bibitem{Boyer1977} Robert S. Boyer and J. Strother Moore, ``A fast
  string searching algorithm'', {\it Commun. ACM}, 20(10), pp.762–772,
  1977.

\bibitem{Zhang2016} Lungan Zhang, Liangxiao Jiang, Chaoqun Li. ``A New
  Feature Selection Approach to Naive Bayes Text Classifiers'' {\it
    International Journal of Pattern Recognition and Artificial
    Intelligence}, 30(9), pp.1650003 2016.

\bibitem{Tang2015} Guoyu Tang, Yunqing Xia, Erik Cambria, Peng Jin,
  Thomas Fang Zheng. ``Document Representation with Statistical Word
  Senses in Cross-Lingual Document Clustering'' {\it International
    Journal of Pattern Recognition and Artificial Intelligence},
  30(9), pp.1559003, 2015.

\bibitem{Tuck2004} N. Tuck, T. Sherwood, B. Calder, and G. Varghese,
  ``Deterministic memory-efficient stringmatching algorithms for
  intrusion detection'', {\it Twenty-third AnnualJoint Conference of
    the IEEE Computer and Communications Societies}, Vol 4,
  pp.2628–2639, 2004.

\bibitem{Bremler2011} A. Bremler-Barr, Y. Harchol, and D. Hay,
  ``Space-time tradeoffs in software-based deep packet inspection'',
  {\it In High Performance Switching and Routing (HPSR), IEEE 12th
    International Conference on}, pp.1–8, 2011.

\bibitem{Ramakrishnan2010} Ramakrishnan Kandhan, Nikhil Teletia, and
  Jignesh M. Patel, ``Sigmatch: Fast and scalable multi-pattern
  matching'', {\it Proceedings of the VLDB Endowment}, 3(1),
  pp.1173–1184, 2010.

\bibitem{Zhou2007} Zongwei Zhou, Yibo Xue, Junda Liu, Wei Zhang, and
  Jun Li, ``MDH: A high speed multi-phase dynamic hash string matching
  algorithm for large-scale pattern set'', {\it Information and
    Communications Security Springer}, pp.201-215, 2007.

\bibitem{Zhan2014} P. Zhan, W. Yuping, and X. Jinfeng, ``An improved
  multi-pattern matching algorithm for large-scale pattern sets'',
  {\it In Computational Intelligence and Security (CIS), Tenth
    International Conference on}, pp.197–200, 2014.

\bibitem{Zhang2009} Baojun Zhang, Xiaoping Chen, Xuezeng Pan, and
  Zhaohui Wu, ``High concurrence wu-manber multiple patterns matching
  algorithm'', {\it Proceedings of the International Symposium on
    Information Proces}, pp.404-409, 2009.

\bibitem{Choi2011} Yoon-Ho Choi, Moon-Young Jung, and Seung-Woo Seo,
  ``A fast pattern matching algorithm with multi-byte search unit for
  highspeed network security'', {\it Computer Communications}, 34(14),
  pp.1750–1763, 2011.

\bibitem{Le2013} Hoang Le and V.K. Prasanna, ``A memory-efficient and
  modular approach for large-scale string pattern matching'', {\it
    Computers, IEEE Transactions on}, 62 (5), pp.844–857, 2013.

\bibitem{Moraru2013} Iulian Moraru and David G. Andersen, ``Exact
  pattern matching with feed-forward bloom filters'', {\it Journal of
    Experimental Algorithmics}, Vol.17, pp.3-4, 2013.

\bibitem{Karp1987} R. M. Karp and M. O. Rabin, ``Efficient randomized
  patternmatching algorithms'', {\it IBM Journal of Research and
    Development}, 31 (2), pp.249-260, 1987.

\bibitem{Agarwal2013} K. Agarwal and R. Polig, ``A high-speed and
  large-scale dictionary matching engine for information extraction
  systems'',{\it Application-Specific Systems, Architectures and
    Processors (ASAP), IEEE 24th International Conference on},
  pp.59–66, 2013.

\bibitem{Zhang2015} Hongli Zhang, Dongliang Xu, Zhihong Tian, and
  Yujian Fan, ``An efficient parallel algorithm for exact
  multi-pattern matching'', {\it Security And Communication Networks},
  8 (9), pp.1688–1697, 2015.

\bibitem{I2015} Tomohiro I, Takaaki Nishimoto, Shunsuke Inenaga, Hideo
  Bannai, and Masayuki Takeda, ``Compressed automata for dictionary
  matching'', {\it Theoretical Computer Science}, Vol.578, pp.30–41,
  2015.

\bibitem{Khancome2013} Chouvalit Khancome and Veera Boonjing, ``A new
  linear-time dynamic dictionary matching algorithm'', {\it Computing
    And Informatics}, 32 (5), pp.897–923, 2013.

\bibitem{Amir2015} Amihood Amir, Avivit Levy, Ely Porat, and B. Riva
  Shalom, ``Dictionary matching with a few gaps'', {\it Theoretical
    Computer Science}, Vol.589, pp.34–46, 2015.

\bibitem{Neuburger2012} Shoshana Neuburger and Dina Sokol. ``Succinct
  2d dictionary matching'', {\it Algorithmica}, 65 (3), pp.662–684,
  2012.

\bibitem{Leis2013} V. Leis, A. Kemper, and T. Neumann. ``The adaptive
  radix tree: Artful indexing for main-memory databases'', {\it In
    Data Engineering(ICDE), IEEE 29th International Conference on},
  pp.38–49, 2013.

\bibitem{Ramakrishna1997} M. V. Ramakrishna and Justin
  Zobel. ``Performance in practice of string hashing functions'', {\it
    In Proceedings of the Fifth International Conference on Database
    Systems for Advanced Applications (DASFAA)}, pp.215–224, 1997.

\bibitem{Kim2015} Kim, HyunJin AND Choi, Kang-Il AND Choi,
  Sang-Il. ``A Memory-Efficient Deterministic Finite Automaton-Based
  Bit-Split String Matching Scheme Using Pattern Uniqueness in Deep
  Packet Inspection'', {\it PLOS ONE}. vol.10, pp.1-24, 2015.

\bibitem{pizzachil} {\it http://pizzachil.dcc.uchile.cl/}

\end{thebibliography}

\vspace*{-0.01in}
%\vspace*{-0.3in}
\noindent
\rule{12.6cm}{.1mm}

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End:

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End: